### 2019-08-01 solver/problem_omega.tex

\begin{table}
    \scriptsize{% (i)
  \begin{center}
    \begin{tabular}{c|cccc||cc}
      \hline
      \multirow{2}{*}{$h/H$} & \multicolumn{4}{c||}{$\mathcal{B}$} & \multicolumn{2}{c}{$\mathcal{B}_{h}$}\\
      \cline{2-7}
      & $\norm{u-u_h}_1$ & $\norm{\lambda-\lambda_h}_0$ & \# & $\kappa$ & \# & $\kappa$\\
      \hline
1        & 8.62E-1(--)  & 1.04E0(--)    &   21 & 7.8  & 24 & 20.9 \\
$2^{-1}$ & 4.4E-1(0.99) & 3.8E-1(1.46)  & 24   & 7.9  & 41 & 21.1 \\
$2^{-2}$ & 2.2E-1(1.00) & 1.4E-1(1.49)  & 22   & 7.9  & 46 & 21.2 \\
$2^{-3}$ & 1.1E-1(1.00) & 4.8E-2(1.50)  & 22   & --   & 47 & --   \\
$2^{-4}$ & 5.5E-2(1.00) & 1.7E-2(1.50)  & 21   & --   & 48 & --   \\
$2^{-5}$ & 2.7E-2(1.00) & 6.0E-3(1.50)  & 21   & --   & 45 & --   \\
\hline
  \end{tabular}
  \end{center}
    }
    \vspace{5pt}
  \scriptsize{%(ii)
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|cccc||cc}
      \hline
      \multirow{2}{*}{$h/H$} & \multicolumn{4}{c||}{$\mathcal{B}$} & \multicolumn{2}{c}{$\mathcal{B}_{h}$}\\
      \cline{2-7}
      & $\norm{u-u_h}_1$ & $\norm{\lambda-\lambda_h}_0$ & \# & $\kappa$ & \# & $\kappa$\\
      \hline
1       & 8.6E-1(--) & 5.4E-1(--)    & 29 & 6.5 & 25 & 16.9\\
$2^{-1}$ & 4.3E-1(0.98) & 2.3E-1(1.20)& 29 & 6.6 & 31 & 17.7\\
$2^{-2}$ & 2.2E-1(1.00) & 1.0E-1(1.19)& 28 & 6.6 & 34 & 18.1\\
$2^{-3}$ & 1.1E-1(1.00) & 4.7E-2(1.13)& 28 & --   & 37 & --   \\
$2^{-4}$ & 5.5E-2(1.00) & 2.2E-2(1.08)& 28 & --   & 37 & --   \\
$2^{-5}$ & 2.7E-2(1.00) & 1.1E-2(1.04)& 27 & --   & 37 & --   \\
\hline
  \end{tabular}
  \end{center}
  \end{minipage}
  }
    \scriptsize{%(iii)
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|cccc||cc}
      \hline
      \multirow{2}{*}{$h/H$} & \multicolumn{4}{c||}{$\mathcal{B}$} & \multicolumn{2}{c}{$\mathcal{B}_{h}$}\\
      \cline{2-7}
      & $\norm{u-u_h}_1$ & $\norm{\lambda-\lambda_h}_0$ & \# & $\kappa$ & \# & $\kappa$\\
      \hline
1       & 8.6E-1(--) & 5.4E-1(--)  & 31 & 6.7 & 21 & 2.5\\
$2^{-1}$ & 4.3E-1(0.98) & 2.3E-1(1.20)& 32 & 6.6 & 21 & 2.5\\
$2^{-2}$ & 2.2E-1(1.00) & 1.0E-1(1.19)& 31 & 6.6 & 21 & 2.5\\
$2^{-3}$ & 1.1E-1(1.00) & 4.7E-2(1.13)& 30 & --  & 20 & --  \\
$2^{-4}$ & 5.5E-2(1.00) & 2.2E-2(1.08)& 28 & --  & 20 & --  \\
$2^{-5}$ & 2.7E-2(1.00) & 1.1E-2(1.04)& 28 & --  & 18 & --  \\
      \hline
  \end{tabular}
  \end{center}
  \end{minipage}
    }
    \vspace{5pt}
    \\
  \scriptsize{%(iv)
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|cccc}
      \hline
      \multirow{2}{*}{$h/H$} & \multicolumn{4}{c|}{$\mathcal{B}$}\\
      \cline{2-5}
      & $\norm{u-u_h}_1$ & $\norm{\lambda-\lambda_h}_0$ & \# & $\kappa$\\
      \hline
1       & 1.7E0(--)    & 2.0E0(--)    & 22 &4.5\\
$2^{-1}$ & 1.0E0(0.71)  & 1.2E0(0.71)  & 27 &5.1\\
$2^{-2}$ & 5.9E-1(0.78) & 7.7E-1(0.69) & 29 &5.4\\
$2^{-3}$ & 3.3E-1(0.85) & 4.9E-1(0.64) & 29 &-- \\
$2^{-4}$ & 1.7E-1(0.91) & 3.3E-1(0.59) & 28 &-- \\
$2^{-5}$ & 9.1E-2(0.95) & 2.2E-1(0.55) & 26 &-- \\
      \hline
  \end{tabular}
  \end{center}
  \end{minipage}
  }
  \scriptsize{%(iv)
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|cccc}
      \hline
      \multirow{2}{*}{$h/H$} & \multicolumn{4}{c|}{$\mathcal{B}$}\\
      \cline{2-5}
      & $\norm{u-u_h}_1$ & $\norm{\lambda-\lambda_h}_0$ & \# & $\kappa$\\
      \hline
1       & 2.0E0(--) & 1.3E0(--)        & 24 & 5.8\\
$2^{-1}$ & 1.4E0(0.53) & 7.0E-1(0.93)   & 29 & 7.1\\
$2^{-2}$ & 9.1E-1(0.62) & 3.5E-1(1.02)  & 34 & 8.0\\
$2^{-3}$ & 5.7E-1(0.68) & 1.6E-1(1.08)  & 36 & -- \\
$2^{-4}$ & 3.5E-1(0.71) & 7.9E-2(1.05)  & 36 & -- \\
$2^{-5}$ & 2.2E-1(0.69) & 4.3E-2(0.89)  & 34 & -- \\
      \hline
  \end{tabular}
  \end{center}
  \end{minipage}
  }
%%%%%%%%%%%%%
  \caption{}
  \label{}
\end{table}