### 2019-08-01 solver/solver.tex

% SIAM Article Template
\documentclass[r]{siamart171218}
% Packages and macros go here
\usepackage[english]{babel}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{epsfig}
\usepackage{float}
\usepackage{fullpage}
\usepackage{color}
\usepackage{enumitem}  
\usepackage{epstopdf}
\usepackage{tikz}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{standalone}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{color}

\usepackage{ifpdf}
\ifpdf
\DeclareGraphicsRule{*}{mps}{*}{}
\fi

\DeclareRobustCommand{\figlegend}[1]{%
  \raisebox{0.5ex}{
    \tikz[]{%
      \fill[very thick, color=#1] (0, 0) rectangle (0.3, 0.1);
    }%
  }
}
% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
%\title{On the weak coupling of 3D and 1D second order elliptic problems}
\title{Notes on robust solvers for systems of coupled multiscale PDEs}

% Authors: full names plus addresses.
\author{
Miroslav Kuchta, Federica Laurino, Kent-Andre Mardal, Paolo Zunino,\thanks{Authors are listed in alphabetical order}
}

\newcommand{\Miro}[1]{\textcolor{cyan}{Miro: #1}}

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  We summarize numerical experiments related to the task of developing
  mesh and parameter indepent solvers for the 3$d$-1$d$ coupled diffusion problems.
  These are personal notes written to keep track of the developments on this
  topic and are to be kept confidential.
\end{abstract}

% REQUIRED
\begin{keywords}
elliptic problems, high dimensionality gap, essential coupling conditions, Lagrange multipliers
\end{keywords}

% REQUIRED
\begin{AMS}
n.a.
\end{AMS}

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

\newcommand{\semi}[1]{\lvert{#1}\rvert}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\jump}[1]{\ensuremath{[\![#1]\!]} }
\newcommand{\avg}[1]{\ensuremath{\left\{\!\left\{#1\right\}\!\right\}} }
\def\ud{u_{\odot}}
\def\vd{v_{\odot}}
\def\ld{\lambda_{\odot}}
\def\md{\mu_{\odot}}
\def\lld{l_{\odot}}
\def\uf{u_{\ominus}}
\def\up{u_{\oplus}}
\def\eps{\epsilon}
\def\nn{\boldsymbol n}
\def\rr{\boldsymbol r}
\def\RR{\boldsymbol R}
\def\kk{\boldsymbol k}
\def\ss{\boldsymbol s}
\def\uu{\boldsymbol u}
\def\vv{\boldsymbol v}
\def\xx{\boldsymbol x}
\def\bu{\overline{u}}
\def\bv{\overline{v}}
\def\tu{\widetilde{u}}
\def\tv{\widetilde{v}}
\def\TT{\boldsymbol T}
\def\NN{\boldsymbol N}
\def\BB{\boldsymbol B}
\def\ttu{\widetilde{\widetilde{u}}}
\def\ttv{\widetilde{\widetilde{v}}}
\def\cv{\check{v}}
\def\mesh{{\cal T}^h}
\def\ball{{\cal B}}
\def\R{\mathbb{R}}
\def\D{\mathcal{D}}
\def\DD{\partial\mathcal{D}}
\def\trace{{\mathcal{T}_\Gamma}}
\def\mtrace{{\overline{\mathcal{T}}_\Lambda}}
\def\ext{\mathcal{E}_\Gamma}
\def\ide{\mathcal{I}}
\def\ii{\hat{\imath}}
\newcommand{\avrd}[1]{\overline{\overline{#1}}}
\newcommand{\avrc}[1]{\overline{#1}}
\newcommand{\refe}[1]{{#1}_{\mathrm{ref}}}

\newcommand{\vertiii}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 
    \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}{Property}[section]
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\section{Introduction}\label{sec:intro}

Let $\langle \cdot , \cdot \rangle_\Lambda$ denote the duality pairing between 
$H^\frac12_{00}(\Lambda)$ and $H^{-\frac12}(\Lambda)$. This note concerns robust
solvers for the problem:
Find $u \in H^1_0(\Omega),\ \ud \in H^1_0(\Lambda), \ \ld \in H^{-\frac12}(\Lambda)$, such that
\begin{subequations}\label{eq:problem2}
\begin{align}
&(u,v)_{H^1(\Omega)} + (\ud,\vd)_{H^1(\Lambda),|\D|} 
+  \langle \mtrace v - \vd, \ld \rangle_{H^{-\frac12}(\Lambda), |\DD|} 
\\
\nonumber
&\qquad\qquad= (f,v)_{L^2(\Omega)} +  (\avrd{g},V)_{L^2(\Lambda),|\D|}
\quad \forall v \in H^1_0(\Omega), \ \vd \in H^1_0(\Lambda)\,,
\\
&   \langle \mtrace u -   \ud, \md \rangle_{H^{-\frac12}(\Lambda),|\DD|} = 0
\quad \forall \md \in H^{-\frac12}(\Lambda)\,,
\end{align}
\end{subequations}
where $\mtrace$ denotes the composition of operators $\avrc{(\cdot)} \circ \trace $.
In particular, we aim at devising algorithms that are robust with respect to
the averaging radius in $\mtrace$.

The (continuous) problem \eqref{eq:problem2} is well posed, in particular
the Lagrange multiplier $\md$ exists in a suitable fractional Sobolev space on
$\Lambda$. Following operator preconditioning framework \cite{mardal2011preconditioning} preconditioners
for the problem rely on fractional Sobolev norms. At the moment the stability
of the discrete problem can be shown using the Lagrange multiplier defined
in the space of piecewise constant functions over the collection of tetrahedra
(of the 3$d$ mesh) that are intersected by $\Lambda$. However, it is not clear
how fractional Sobolev norms on such a space should be constructed (in contrast
to the case where the multiplier is defined on $\Lambda$). We shall therefore
establish the preconditioners for the \emph{discrete} problem using different
norms than those used in the analysis of the \emph{continuous} one.

As a stepping stone to \eqref{eq:problem2} we shall consider a related
2$d$-1$d$ problem\\
%Stong
\begin{minipage}{0.57\textwidth}
  \begin{equation}\label{eq:coupled}
    \begin{aligned}
      -\Delta u^i &= f &\mbox{ in }\Omega^{i}, i\in\left\{+, -\right\},\\
      \jump{u} &= 0 &\mbox{ on }\Gamma,\\
      -\Delta \hat{u} +\hat{u} - \jump{\nabla u\cdot n}&= \hat{f} &\mbox{ on }\Gamma,\\
      \epsilon u - \hat{u} &= h&\mbox{ on }\Gamma.\\
    \end{aligned}
\end{equation}
\null
\par\xdef\tpd{\the\prevdepth}
\end{minipage}
\hfill
\begin{minipage}{0.37\textwidth}
  \begin{figure}[H]
    \begin{center}
\begin{tikzpicture}
  \node[black, fill=white, opacity=0.9] at (1, 2) {$\Omega^{-}$};
  \node[black, fill=white, opacity=0.9] at (3, 2) {$\Omega^{+}$};

  % Boundaries
  \draw[red!60!white] (0, 0) -- (0, 4);
  \draw[red!60!white] (4, 0) -- (4, 4);
  \node[red!60!white, opacity=0.9, right] at (4, 1) {$\Gamma_D$};

  \draw[blue!60!white] (0, 0) -- (4, 0);
  \draw[blue!60!white] (0, 4) -- (4, 4);
  \node[blue!60!white, opacity=0.9, below] at (3, 0) {$\Gamma_N$};

  \draw[black, very thick] (2, 0) -- (2, 4);
  \draw[->, black] (2, 1) -- (3, 1);
  \node[fill=white, opacity=0.9, above] at (2.5, 1) {$n$};
  \node[fill=white, opacity=0.9, right] at (2.025, 2) {$\Gamma$};
\end{tikzpicture}
    \end{center}
    %\vspace{-20pt}
    \caption{Schematic domain of coupled multiscale problem.}
    \label{fig:coupled_domain}
    %\vspace{5pt}    
\end{figure}
\end{minipage}
%
Here \eqref{eq:coupled} shall be equipped with Dirichlet boundary conditions
(for $u^i$) on $\Gamma_D$, while Neumann conditions are used on $\Gamma_N$
(for both the unknowns). Note that the role of constant parameter $\epsilon$ is similar
to that of averaging radius in \eqref{eq:problem2}.

Let $\Omega=\Omega^{+}\cup\Omega^{-}$. Introducing a Lagrange multiplier
$p=\epsilon^{-1}\jump{\nabla u \cdot n}\in Q=H^{-1/2}(\Gamma)$
the weak form of \eqref{eq:coupled} reads: Find $u\in V=H^1_{0, \Gamma_D}(\Omega)$,
$\hat{v}\in\hat{V}=H^1(\Gamma)$, $p\in Q$ such that
%
\begin{equation}\label{eq:coupled_weak}
  \begin{aligned}
    &\int_{\Omega}\nabla u\cdot \nabla v &\phantom{+\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &+\epsilon\int_{\Gamma} p v &= \int_{\Omega}f v \quad &\forall v\in V,\\
    &\phantom{\int_{\Omega}\nabla u\cdot \nabla v} &+{\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v} + \hat{u}\hat{v}} &-\int_{\Gamma} p\hat{v} &= \int_{\Gamma}\hat{f}\hat{v} \quad &\forall \hat{v}\in \hat{V},\\
    &\epsilon\int_{\Gamma}u q &-\int_{\Gamma}\hat{u}q &\phantom{-\int{\Gamma}q\hat{v}} &= \int_{\Gamma}h q \quad &\forall q\in Q.
  \end{aligned}
\end{equation}

In the following it will be convenient to work with the operator form of
\eqref{eq:coupled_weak}. In particular the left-hand-side defines an operator
%
\begin{equation}\label{eq:operator2d}
  \mathcal{A} = \begin{pmatrix}
    -\Delta_{\Omega} & & -\epsilon T^{\prime}\\
    & -\Delta_{\Gamma} + I & I\\
    -\epsilon T & I & \\
  \end{pmatrix} =
  \begin{pmatrix}
    A_2 &  & B_2'\\
    & A_1 & -B_1'\\
    B_2 & -B_1 & \\
   \end{pmatrix}.
\end{equation}
In \cite{kuchta2016preconditioners} it is shown that $\mathcal{A}$ is an
isomorphism from $W$ to its dual space where
\[
W=H^1(\Omega)_{0, \Gamma_D}\times H^1{(\Gamma)}\times(\epsilon H^{-1/2}(\Gamma)\cap H^{-1}(\Gamma)).
\]
It is further shown that discretization in terms of continuous linear Lagrange
elements (P1) for all the unknowns is inf-sup stable and in turn the diagonal operator
\begin{equation}\label{eq:P1_precond}
\mathcal{B}=\text{diag}(-\Delta_{\Omega}, -\Delta_{\Gamma}+I, \epsilon^2(-\Delta_{\Gamma})^{-1/2}+(-\Delta_{\Gamma}+I)^{-1})^{-1}
\end{equation}
is a robust preconditioner for $\mathcal{A}$. For the sake of completness
(and verifying the implementation) Table \ref{tab:P1} shows the number of
MinRes iterations and condition numbers of the preconditioned problem for
different levels of refinements and values of the parameter $\epsilon$. The
stability is evident.
%
\begin{table}
    \footnotesize{
      \begin{minipage}{0.49\textwidth}
          \begin{center}
            \begin{tabular}{c|cccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{4}{c}{$h$}\\
              \cline{2-5}
 &   $2^{-2}$  & $2^{-3}$ & $2^{-4}$ & $2^{-5}$   \\  
\hline
$10^{6}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{4}$ &	7.73&	7.82&	7.86&	7.87\\
$10^{2}$   &	7.73&	7.82&	7.86&	7.87\\
1     &	7.39&	7.79&	8.01&	8.13\\
$10^{-2}$  &	2.62&	2.63&	2.63&	2.65\\
$10^{-4}$ &	2.62&	2.62&	2.62&	2.62\\
$10^{-6}$ &	2.62&	2.62&	2.62&	2.62\\
\hline
\end{tabular}
\end{center}
      \end{minipage}
    }
          \footnotesize{
            \begin{minipage}{0.49\textwidth}
              \begin{center}
            \begin{tabular}{c|cccccc} \hline
              \multirow{2}{*}{$\epsilon$} & \multicolumn{6}{c}{$h$}\\
              \cline{2-7}
 &   $2^{-3}$ & $2^{-4}$ & $2^{-5}$ & $2^{-6}$ & $2^{-7}$ & $2^{-8}$  \\  
              \hline
$10^{6}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{4}$ & 21 & 22 & 22 & 21 & 21 & 19\\
$10^{2}$ & 21 & 22 & 22 & 21 & 21 & 19\\
1 & 24 & 28 & 26 & 26 & 27 & 25\\
$10^{-2}$ & 10 & 10 & 10 & 12 & 13 & 14\\
$10^{-4}$ & 6 & 6 & 6 & 6 & 7 & 7\\
$10^{-6}$ & 4 & 4 & 4 & 4 & 4 & 4\\
\hline
\end{tabular}                
              \end{center}
        \end{minipage}
      }
          \caption{
            Problem \eqref{eq:coupled_weak} discretized with P1 elements.
            Condition number (left) and iterations counts (right) using
            preconditioner \eqref{eq:P1_precond}.
          }
          \label{tab:P1} 
%\label{table1}
\end{table}

Preconditioner \eqref{eq:P1_precond} was derived under the assumption that
the finite element mesh $\Gamma_h$ of $\Gamma$ consists of facets of the finite
element mesh $\Omega_h$ of $\Omega$, see also Figure \ref{fig:domains} (i). This
assumption is too limiting; for practical purposes the meshes should be independent
and a more common case is that the trace mesh of $\Omega_h$ on $\Gamma$ is coarser
than $\Gamma_h$. In the following we shall pursue formulations (and corresponding
solvers) of \eqref{eq:coupled_weak} which allow for different meshes and resolutions.
Our ultimate goal is a stable formulation in which the multiplier is not defined
on $\Gamma_h$ but rather on cells of $\Omega_h$ intersected by the curve, cf. Figure
\ref{fig:domains} (iv) and (v).
%
\begin{figure}
  \begin{center}
    \includegraphics[width=0.4\textwidth]{./img/conform.tex}\\

  \includegraphics[width=0.4\textwidth]{./img/conform_fine.tex}\hfill
  \includegraphics[width=0.4\textwidth]{./img/cut_fine.tex}\\
  \includegraphics[width=0.4\textwidth]{./img/nocut_lm2d.tex}\hfill
    \includegraphics[width=0.4\textwidth]{./img/cut_lm2d.tex}
  \end{center}
  \caption{Schematic geometric setting and discretization of mutliscale problems.
    Position of interface $\Gamma$ is denoted by \figlegend{green!50!black}. Meshes
    underlying discrete approximation spaces $\hat{V}_h$ and $Q_h$ of $\hat{V}$ and $Q$ are denoted respectively
    by \figlegend{white!50!blue} and \figlegend{white!50!red}. Geometric correspondence
    is signified by dashed line. Cases: (i) trace mesh of $V_h$ is used for $\hat{V}_h$ and $Q_h$,
    (ii) $\Gamma$ does not intersect $\Omega_h $ cell interiors, (iii) $\Gamma$ does
    intersect cell interiors, (iv) as (ii) with $Q_h$ defined on intersected cells and
    (v) as (iii) with $Q_h$ defined on intersected cells.
  }
 \label{fig:domains}
\end{figure}

The rest of the report is structured as follows. Considering \eqref{eq:operator2d}
we first treat the problems related to domains $\Omega$, i.e. $\mathcal{A}_2=\left[A_2, B_2'; B_2, 0\right]$,
and $\Gamma$, i.e. $\mathcal{A}_1=\left[A_1, B_1'; B_1, 0\right]$, individually\footnote{This is done for
clarity of exposition, to better fix the ideas and debugging}. More precisely, in
\S\ref{sec:problem_omega} a set of preconditioners for different formulations of the Babu{\v s}ka
problem is presented culminating in a formulation having the discrete Lagrange multiplier on the
cut cells. In \S\ref{sec:problem_gamma} we deal with instabities in $\mathcal{A}_1$ which arise
due to choices of discretization of the Lagrange multiplier space in $\mathcal{A}_2$. Finally in
\S\ref{sec:coupled} the insights from individual problems are combined to yield robust solvers
for the coupled problem.

\section{Coupling 2$d$-1$d$}\label{sec:problem_omega}
\input{problem_omega}

\section{Coupling 1$d$-1$d$}\label{sec:problem_gamma}
\input{problem_gamma}

\section{Coupling 2$d$-1$d$-1$d$}\label{sec:coupled}
\input{coupled}

\bibliographystyle{siamplain}
\bibliography{solver}
\end{document}