### 2020-01-09 lm3d1d_vs2/3d1d_coupled_LM.tex

% SIAM Article Template
\documentclass[r]{siamart171218}
% Packages and macros go here
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{float}
\usepackage{fullpage}
\usepackage{color}
\usepackage{enumitem}  
\usepackage{epstopdf}
\usepackage{stmaryrd}
\usepackage{tikz, subfigure}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
%\title{On the weak coupling of 3D and 1D second order elliptic problems}
\title{Coupling PDEs on 3D-1D domains with Lagrange multipliers}

% Authors: full names plus addresses.
\author{Miroslav Kuchta, Federica Laurino, Kent-Andre Mardal, Paolo Zunino,\thanks{Authors are listed in alphabetical order}}

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
These are personal notes written to keep track of the developments on this topic, to be kept confidential.
\end{abstract}

% REQUIRED
\begin{keywords}
elliptic problems, high dimensionality gap, essential coupling conditions, Lagrange multipliers
\end{keywords}

% REQUIRED
\begin{AMS}
n.a.
\end{AMS}

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

% definitions here

\def\ud{u_{\odot}}
\def\vd{v_{\odot}}
\def\ld{\lambda_{\odot}}
\def\md{\mu_{\odot}}
\def\lld{l_{\odot}}
\def\uf{u_{\ominus}}
\def\up{u_{\oplus}}
\def\eps{\epsilon}
\def\nn{\boldsymbol n}
\def\rr{\boldsymbol r}
\def\RR{\boldsymbol R}
\def\kk{\boldsymbol k}
\def\ss{\boldsymbol s}
\def\uu{\boldsymbol u}
\def\vv{\boldsymbol v}
\def\xx{\boldsymbol x}
\def\bu{\overline{u}}
\def\bv{\overline{v}}
\def\tu{\widetilde{u}}
\def\tv{\widetilde{v}}
\def\TT{\boldsymbol T}
\def\NN{\boldsymbol N}
\def\BB{\boldsymbol B}
\def\ttu{\widetilde{\widetilde{u}}}
\def\ttv{\widetilde{\widetilde{v}}}
\def\cv{\check{v}}
\def\mesh{{\cal T}^h}
\def\ball{{\cal B}}
\def\R{\mathbb{R}}
\def\D{\mathcal{D}}
\def\DD{\partial\mathcal{D}}
\def\trace{{\mathcal{T}_\Gamma}}
\def\mtrace{{\overline{\mathcal{T}}_\Lambda}}
\def\ext{\mathcal{E}_\Gamma}
\def\ide{\mathcal{I}}
\def\ii{\hat{\imath}}
\newcommand{\avrd}[1]{\overline{\overline{#1}}}
\newcommand{\avrc}[1]{\overline{#1}}
\newcommand{\refe}[1]{{#1}_{\mathrm{ref}}}

\newcommand{\vertiii}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 
    \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}{Property}[section]
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\section{Introduction}\label{sec:intro}

We address the geometrical configuration of the problem for a 3D coupled problem formulation based on from Dirichlet-Neumann interface conditions. Then, we apply a model reduction technique that transforms the problem into 3D-1D coupled PDEs. We develop and analyze a robust definition of the coupling operators form a 3D domain, $\Omega$, to 1D manifold, $\Lambda$, and vice versa. This is a non trivial objective because the standard trace operator form a domain $\Omega$ to a subset $\Lambda$ is not well posed if $\Lambda$ is a manifold of co-dimension two of $\Omega$.

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\section{Problem setting}\label{sec:setting}

Let $\Omega \subset \mathbb{R}^3$ be a bounded, convex open set. Let $\Omega_\ominus$ be a generalized cylinder embedded into $\Omega$ and be $\Omega_\oplus = \Omega \setminus \overline{\Omega}_\ominus$ be the complementary set of the cylinder. We also introduce the set $\Lambda$, a 1D manifold that is the centerline of $\Omega_\ominus$. We define the arc-length coordinate along $\Lambda$, denoted by $s \in (0,S)$. We denote with $\mathcal{D}(s)$ and $\partial\mathcal{D}(s)$ a cross section of $\Omega_\ominus$ and its boundary, respectively. 
%In what follows, we assume for simplicity of notation that $\Sigma$ has a constant cross section, but this is not a restriction of the approach. 
We also assume that $\Omega_\ominus$ crosses $\Omega$ from side to side and we call $\Gamma$ the lateral (cylindrical) surface of $\Omega_\ominus$, while the upper and lower side faces of $\Omega_\ominus$ belong to $\partial\Omega$. We refer to Figure \ref{fig1} for an illustration of the notation.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{3D-1D-simple.png}
\end{center}
\caption{Geometrical setting of the problem}
\label{fig1}
\end{figure}

We consider the problem arising from \emph{Dirichlet-Neumann} conditions. 
It consists to find $\up,\uf$ s.t.:
\begin{subequations}\label{eq:dirneu}
\begin{align}
\label{eq:dirneu1}
- \Delta \up  + \up &= f  && \text{ in } \Omega_{\oplus},\\
\label{eq:dirneu2}
- \Delta \uf  + \uf &= g  && \text{ in } \Omega_\ominus,\\
\label{eq:dirneu3}
-\nabla \uf \cdot \nn_{\ominus} &= -\nabla \up \cdot \nn_{\ominus}  && \text{ on } \Gamma,\\
\label{eq:dirneu4}
\uf &= \up && \text{ on }  \Gamma,\\
\label{eq:dirneu5}
\up &= 0 && \text{ on } \partial \Omega\,.
\end{align}
\end{subequations}

The objective of this work is to derive and analyze a simplified version of problem \eqref{eq:dirneu}, where the domain $\Omega_\ominus$ shrinks to its centerline $\Lambda$ and the corresponding partial differential equation is averaged on the cylinder cross section, namely $\mathcal{D}$. This new problem setting will be called the \emph{reduced} problem. Form the mathematical standpoint it is more challenging than \eqref{eq:dirneu}, because it involves the coupling of 3D/1D elliptic problems.
For the model reduction process, we decompose integrals as follows, for any sufficiently regular function $w$,
\begin{equation*}
\int_{\Omega_\ominus} w d\omega 
= \int_\Lambda \int_{\mathcal{D}} w d\sigma ds
= \int_\Lambda |\mathcal{D}|\avrd{w} ds\,,
\quad
\int_{\Gamma} w d\sigma 
= \int_\Lambda \int_{\partial\mathcal{D}} w d\gamma ds
= \int_\Lambda  |\partial \mathcal{D}| \avrc{w} ds\,,
\end{equation*}
where $\avrd{w}, \ \avrc{w}$ denote the following mean values respectively,
\begin{equation*}
\avrd{w} = |\mathcal{D}|^{-1} \int_{\mathcal{D}} w d\sigma\,,
\quad
\avrc{w} = |\partial\mathcal{D}|^{-1} \int_{\partial\mathcal{D}} w d\gamma\,.
\end{equation*}

We apply the model reduction approach at the level of the variational formulation.
We start from the variational formulation of problem \eqref{eq:dirneu}, 
that is to find $\up \in H^1_{\partial\Omega}(\Omega_\oplus), \ \uf \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus), \ \lambda \in H^{-\frac12}(\partial\Omega_\ominus)$ s.t.
\begin{subequations}\label{eq:weak_dirneu}
\begin{align}
&(u_\oplus,v_\oplus)_{H^1(\Omega_\oplus)} + (u_\ominus,v_\ominus)_{H^1(\Omega_\ominus)} 
+ \langle  v_\oplus - v_\ominus, \lambda \rangle_{H^{-\frac12}(\Gamma)} 
\\
\nonumber
&\qquad\qquad = (f,v_\oplus)_{L^2(\Omega_\oplus)} + (g,v_\ominus)_{L^2(\Omega_\ominus)}
\quad \forall v_\oplus \in H^1_{\partial\Omega}(\Omega_\oplus), \ v_\ominus \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus)
\\
& \langle u_\oplus - u_\ominus, \mu \rangle_{H^{-\frac12}(\Gamma)} = 0
\quad \forall  \mu \in H^{-\frac12}(\Gamma)\,,
\end{align}
\end{subequations}
where $\langle v, \mu \rangle_{H^{-\frac12}(\Gamma)}$ denotes the duality pairing between 
$ \mu \in H^{-\frac12}(\Gamma)$ and $v \in H^{\frac12}(\Gamma)$.
In this case, the additional variable $\lambda$ is equivalent to $\lambda  =  \nabla \uf \cdot \nn_\ominus$.

Using a model reduction approach based on averaging, we end up with two different formulations of a reduced problem
for the unknown $u$ defined on the entire 3D domain $\Omega$, coupled with the unknown $\ud$, defined on the 1D manifold $\Lambda$
and a Lagrange multiplier defined either on $\Gamma$ (problem 1) or on $\Lambda$ (problem 2). 
The scope of this work is to compare them, with the aim to determine which is the most suitable to set up a computational model for 3D-1D coupled PDEs.

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\subsection{Topological model reduction}
\subsubsection*{Model reduction of the problem on $\Omega_{\ominus}$}

We apply the averaging technique to equation \eqref{eq:dirneu2}. In particular, we consider an arbitrary portion $\mathcal{P}$ of the cylinder $\Omega_\ominus$, with lateral surface $\Gamma _{\mathcal{P}}$ and bounded by two perpendicular sections to $\Lambda$, namely $\mathcal{D}(s_1), \ \mathcal{D}(s_2)$ with $s_1<s_2$. We have,
\begin{multline*}
\int_{\mathcal{P}} -\Delta \uf + \uf d\omega =
-\int_{\partial \mathcal{P}} \nabla \uf \cdot \nn_{\ominus} \, d\sigma  + \int_{\mathcal{P}}\uf d\omega=
\\
 \int_{\mathcal{D}(s_1)} \partial_s \uf d\sigma -  \int_{\mathcal{D}(s_2)} \partial_s \uf d\sigma -  \int_{\Gamma_{\mathcal{P}}} \nabla \uf \cdot \nn_{\ominus} d\sigma +\int_{\mathcal{P}}\uf d\omega
\end{multline*}
By the fundamental theorem of integral calculus combined with the Reynolds transport Theorem, 
being $\nu$ the normal deformation of the boundary along $(0,S)$, we have,
\begin{equation*}
\begin{split}
\int_{\mathcal{D}(s_1)} \partial_s \uf d\sigma -  \int_{\mathcal{D}(s_2)} \partial_s \uf d\sigma 
&= -\int_{s_1}^{s_2} d_s \int_{\mathcal{D}(s)}  \partial_s \uf d\sigma ds
\\
&= -\int_{s_1}^{s_2} d_{ss}^2 \int_{\D(s)} u_{\ominus} \, d\sigma \, ds +  \int_{s_1}^{s_2} d_s \left( \int_{\DD(s)} \nu u_{\ominus} \, d\gamma \right) ds\,,
\end{split}
\end{equation*}
and assuming that $\D(s)$  can not change shape, we have
\begin{equation*}
\begin{split}
-\int_{s_1}^{s_2} d_{ss}^2 \int_{\D(s)} u_{\ominus} \, d\sigma \, ds + \int_{s_1}^{s_2} d_s \left( \int_{\DD(s)} \nu u_{\ominus} \, d\gamma \right)\, ds
&=-\int_{s_1}^{s_2} \left[d_{ss}^2 (|\D(s)|\avrd{u}_{\ominus}) -  d_s(  \nu |\DD(s)| \avrc{u}_{\ominus} ) \right] \, ds
\\
&=-\int_{s_1}^{s_2} \left[d_{ss}^2 (|\D(s)|\avrd{u}_{\ominus}) -  d_s\left(d_s(|\D(s)|) \avrc{u}_{\ominus} \right) \right] \, ds.
\end{split}
\end{equation*}
Moreover, we have
\begin{equation*}
\int_{\Gamma_{\mathcal{P}}} \nabla \uf \cdot \nn_{\ominus} d\sigma =  \int_{\Gamma_{\mathcal{P}}} \lambda \, d\sigma
=  \int_{s_1}^{s_2} \int_{\partial\mathcal{D}(s)} \lambda d\gamma \, ds 
= \int_{s_1}^{s_2} |\DD| \avrc{\lambda} \,ds\,.
\end{equation*}
From the combination of all the above terms with the right hand side, we obtain that the solution $\uf$ of \eqref{eq:dirneu2} satisfies,
\begin{equation*}
\int_{s_1}^{s_2} \left[ 
-d_{ss}^2 (|\D(s)|\avrd{u}_{\ominus}) +  d_s\left(d_s(|\D(s)|) \avrc{u}_{\ominus}\right) + |\D(s)|\avrd{u}_{\ominus} 
- |\partial\mathcal{D}(s)| \avrc{\lambda} 
\right] \, ds = \int_{s_1}^{s_2} |\mathcal{D}(s)| \avrd{g}\, ds \,.
\end{equation*}
Since the choice of the points $s_1,s_2$ is arbitrary, we conclude that the following equation holds true,
\begin{equation}\label{eq:averaged_f}
-d_{ss}^2 (|\D(s)|\avrd{u}_{\ominus}) +  d_s\left(d_s(|\D(s)|) \avrc{u}_{\ominus}\right) + |\D(s)|\avrd{u}_{\ominus} 
- |\partial\mathcal{D}(s)| \avrc{\lambda}
= |\mathcal{D}(s)| \avrd{g} \quad \text{on} \ \Lambda\,,
\end{equation}
which is complemented by the following conditions at the boundary of $\Lambda$,
\begin{equation}\label{eq:averaged_f_boundary}
|\D(s)| d_s \avrd{u}_{\ominus} = 0, \quad d_s |\D(s)| = 0, \quad \text{on} \quad s=0,S.
\end{equation}
Then, we consider variational formulation of the averaged equation \eqref{eq:averaged_f}.
After multiplication by a test function $\vd \in H^1(\Lambda)$, integration on $\Lambda$ and suitable application of integration by parts, we obtain,
\begin{multline*}
\int_\Lambda d_s (|\D(s)| \avrd{u}_{\ominus} ) d_s \vd \, ds - d_s (|\D(s)| \avrd{u}_{\ominus} ) \vd |_{s=0}^{s=S}
- \int_\Lambda (d_s |\D(s)|)\avrc{u}_{\ominus} d_s \vd \, ds + (d_s |\D(s)|) \avrc{u}_{\ominus} \vd |_{s=0}^{s=S}
\\
+\int_\Lambda |\D(s)| \avrd{u_\ominus} \vd - \int_\Lambda |\DD(s)| \avrc{\lambda} \vd \, ds
= \int_\Lambda |\D(s)| \avrd{g} V\, ds\,.
\end{multline*}
Using boundary conditions, 
the identity $d_s (|\D(s)| \avrd{u}_{\ominus} ) = |\D(s)| d_s \avrd{u}_{\ominus} + d_s (|\D(s)|) \avrd{u}_{\ominus}$
and reminding that $d_s |\D(s)|)/|\DD(s)| = \nu$,
we obtain,
\begin{equation}\label{eq:averaged_f_weak}
(d_s \avrd{u}_{\ominus}, d_s \vd )_{\Lambda,|\D|} 
+ ( \nu (\avrd{u}_{\ominus} - \avrc{u}_{\ominus}), d_s \vd)_{\Lambda,|\DD|} +(\avrd{u_\ominus}, \vd)_{\Lambda, |\D|}
- (\avrc{\lambda},\vd)_{\Lambda,|\DD|}
= (\avrd{g},V)_{\Lambda,|\D|}\,.
\end{equation}
where we have introduced the following weighted inner product notation,
\begin{equation*}
(\ud,\vd)_{\Lambda,w} = \int_0^S w(s) \ud(s) \vd(s) ds\,.
\end{equation*}
Let us now formulate the modelling assumption that allows us to reduce equation \eqref{eq:averaged_f_weak} to a solvable one-dimensional (1D) model.
More precisely, we assume that:
\begin{description}
\item[A1] the function $\uf$ has a \emph{uniform profile} on each cross section $\mathcal{D}(s)$, namely $\uf(r,s,t) = \ud(s)$.
\end{description}
Therefore, observing that $\ud=\avrc{u}_{\ominus}=\avrd{u}_{\ominus}$, problem \eqref{eq:averaged_f_weak} consists to find $\ud \in H^1(\Lambda)$ such that
\begin{equation}\label{eq:1Ddirneu_weak}
 (d_s \ud, d_s \vd)_{\Lambda,|\mathcal{D}|} 
+(\ud, \vd)_{\Lambda, |\D|}
-(\avrc{\lambda},\vd)_{\Lambda, |\DD|}
=  (\avrd{g},\vd)_{\Lambda, |\mathcal{D}|}
\quad \forall \vd \in H^1(\Lambda)\,.
\end{equation}

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\subsubsection*{Topological model reduction of the problem on $\Omega_{\oplus}$}
We focus here on the subproblem of \eqref{eq:dirneu1} related to $\Omega_{\oplus}$.
We multiply both sides of \eqref{eq:dirneu1} by a test function $v\in H^1_0(\Omega)$ and integrate on $\Omega_\oplus$. Integrating by parts and using boundary and interface conditions, we obtain
\begin{equation*}
\begin{split}
\int_{\Omega _{\oplus}}fv\,d\omega&=
\int _{\Omega _{\oplus}}\nabla \up \cdot\nabla v\,d\omega -\int _{\partial \Omega _{\oplus}}\nabla \up \cdot \nn_{\oplus} v\,d\sigma + \int_{\Omega_{\oplus}} \up v
\\
&=\int _{\Omega _{\oplus}}\nabla \up \cdot\nabla v\,d\Omega -\int_{\Gamma} \nabla \up \cdot \nn_{\oplus} v + \int_{\Omega_{\oplus}} \up v
\\
&=\int _{\Omega _{\oplus}}\nabla \up \cdot\nabla v\,d\Omega +\int_{\Gamma} \lambda v + \int_{\Omega_{\oplus}} \up v.
\end{split}
\end{equation*} 
Then, we make the following modelling assumptions:
\begin{description}
\item[A2] we identify the domain $\Omega_{\oplus}$ with the entire $\Omega$, 
and we correspondingly omit the subscript $\oplus$ to the functions defined on $\Omega_{\oplus}$,
namely
\begin{equation*}
\int_{\Omega_{\oplus}} u_\oplus\, d\omega \simeq \int_{\Omega} u\, d\omega\,.
\end{equation*}
\end{description}
Therefore, we obtain 
\begin{equation*}
(\nabla u ,\nabla v)_{\Omega} +(u,v)_{\Omega}+(\lambda, v)_{\Gamma}  =(f,v)_{\Omega}
\end{equation*}
and combining with \eqref{eq:1Ddirneu_weak} we obtain the first formulation of the reduced problem.

\subsubsection*{Problem 1 (3D-1D-3D)}
Let $\langle \cdot , \cdot \rangle_\Gamma$ denote the duality pairing between 
$H^\frac12_{00}(\Gamma)$ and $H^{-\frac12}(\Gamma)$. The problem consists to find $u \in H^1_0(\Omega),\ \ud \in H^1_0(\Lambda), \ \lambda \in H^{-\frac12}(\Gamma )$, such that
\begin{subequations}\label{eq:problem1}
\begin{align}
&(u,v)_{H^1(\Omega)} + (\ud,\vd)_{H^1(\Lambda),|\D|} 
+ \langle \trace v  - \ext \vd, \lambda \rangle_\Gamma
\\
\nonumber
&\qquad\qquad= (f,v)_{L^2(\Omega)} +  (\avrd{g},\vd)_{L^2(\Lambda),|\D|}
\quad \forall v \in H^1_0(\Omega), \ \vd \in H^1(\Lambda)
\\
&   \langle \trace u - \ext \ud , \mu \rangle_\Gamma = 0
\quad \forall \mu \in H^{-\frac12}(\Gamma)\,.
\end{align}
\end{subequations}
Here, $\trace:H^1_0(\Omega) \rightarrow H^{\frac 12}_{00}(\Gamma)$ denotes the trace operator from $\Omega$ to $\Gamma$ and $\ext: H^1_0(\Lambda) \rightarrow H^1_{0}(\Gamma)$ denotes the uniform extension from $\Lambda$ to $\Gamma$.

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
Now, we apply a topological model reduction of the interface conditions, namely we go from a 3D-1D-2D formulation
involving sub-problems on $\Omega$ and $\Lambda$ and coupling operators defined on $\Gamma$
to a 3D-1D-1D formulation where the coupling terms are set on $\Lambda$. 
To this purpose, let us write the Lagrange multiplier and the test functions on every cross section $\partial\mathcal{D}(s)$ as their average plus some fluctuation,
\begin{equation*}
\lambda=\avrc{\lambda}+\tilde{\lambda}, \qquad v=\avrc{v}+\tilde{v},
\quad \text{on} \ \partial\mathcal{D}(s)\,,
\end{equation*}
where $\avrc{\tilde{\lambda}}=\avrc{\tilde{v}}=0$. 
Therefore, the coupling term on $\Gamma$ can be decomposed as,
\begin{equation*}
\int_{\Gamma}\lambda v\, d\sigma
=\int _{\Lambda}  \int_{\partial\mathcal{D}(s)} (\avrc{\lambda}+\tilde{\lambda})(\avrc{v}+\tilde{v})d\gamma ds
= \int_{\Lambda}|\partial\mathcal{D}(s)| \avrc{\lambda}\avrc{v}\,ds+\int_{\Lambda}  \int_{\partial\mathcal{D}(s)} \tilde{\lambda}\tilde{v}d\gamma ds\,.
\end{equation*}

Then, we make the following modelling assumptions:
\begin{description}
\item[A3] we assume that the product of fluctuations is small, namely
\begin{equation*}
\int_{\partial\mathcal{D}(s)} \tilde{\lambda}\tilde{v} d\gamma \simeq 0\,
\end{equation*}
\end{description}
and we obtain 
\begin{equation*}
(\nabla u ,\nabla v)_{\Omega} +(u,v)_{\Omega}+(\avrc \lambda, \avrc v)_{\Lambda , |\DD|}  =(f,v)_{\Omega},
\end{equation*}
which, combined with \eqref{eq:1Ddirneu_weak} leads to the second formulation of the reduced problem.

\subsection{Problem 2 (3D-1D-1D)}
Let $\langle \cdot , \cdot \rangle_\Lambda$ denote the duality pairing between 
$H^\frac12_{00}(\Lambda)$ and $H^{-\frac12}(\Lambda)$.
The problem requires to find $u \in H^1_0(\Omega),\ \ud \in H^1_0(\Lambda), \ \ld \in H^{-\frac12}(\Lambda)$, such that
\begin{subequations}\label{eq:problem2}
\begin{align}
&(u,v)_{H^1(\Omega)} + (\ud,\vd)_{H^1(\Lambda),|\D|} 
+  \langle \mtrace v - \vd, \ld \rangle_{H^{-\frac12}(\Lambda), |\DD|} 
\\
\nonumber
&\qquad\qquad= (f,v)_{L^2(\Omega)} +  (\avrd{g},V)_{L^2(\Lambda),|\D|}
\quad \forall v \in H^1_0(\Omega), \ \vd \in H^1_0(\Lambda)\,,
\\
&   \langle \mtrace u -   \ud, \md \rangle_{H^{-\frac12}(\Lambda),|\DD|} = 0
\quad \forall \md \in H^{-\frac12}(\Lambda)\,,
\end{align}
\end{subequations}
where $\mtrace$ denotes the composition of operators $\trace \circ \avrc{(\cdot)}$.
We notice that all the integrals of the reduced problem are well defined because 
$\mtrace: H^1_0(\Omega) \rightarrow H^\frac12_{00}(\Lambda)$ as shown in the following lemma.

%\begin{lemma}\label{lemma:H12norm}
%{\color{red}TO DO: generalize to not uniform $\DD$} 
%When $\Gamma$ is a cylinder, if $u\in H_{00}^{\frac 12}(\Gamma)$, then $\avrc{u}\in H_{00}^{\frac 12}(\Lambda)$. Moreover, if $u\in H^{\frac 12}_{00}(\Gamma)$ is constant on each cross section, namely $u(s,\theta)=u(s)$, then 
%\begin{equation*}
%\|u\|_{H^{\frac 12}_{00}(\Gamma)}=2\pi R \|u\|_{H^{\frac 12}_{00}(\Lambda)}.
%\end{equation*}
%\end{lemma}
%\begin{proof}
%Let us denote as $\phi _{ij}$ and $\rho _{ij}$, for $i=1,2,\dots$, $j=0,1,\dots$, the eigenfunctions and the eigenvalues of the laplacian on $\Gamma$, and with $\phi _i$ and $\rho _i$ the eigenfunctions and the eigenvalues of the laplacian on $\Lambda$. In particular,
%\begin{align*}
%\phi _{ij}(s,\theta)=sin (i\pi s)\left( cos(j\theta)+ sin(j\theta) \right),\\
%\rho_{ij}=i\pi ^2+\frac{j^2}{R^2},\\
%\phi _{i}(s)=sin (i\pi s),\\
%\rho _i = i\pi ^2.
%\end{align*}
%It is easy to verify that 
%\begin{eqnarray}
%\label{null_int_eigenf}
%\int_0^{2\pi} \phi _{ij}(s,\theta)=0 \quad \forall j>0, \forall i \\
%\label{nonull_int_eigenf}
%\int_0^{2\pi} \phi _{ij}(s,\theta)= 2\pi R \, \sin(i \pi s) \quad \mbox{if } j=0, \forall i  .\\
%\end{eqnarray}
%Moreover we recall that $\phi_{i,j}(s,\theta)$ and $\phi _i(s)$ are orthogonal basis of $L^2(\Gamma)$ and $L^2(\Lambda)$ respectively. Therefore,
%\begin{multline*}
%\avrc{u}(s)=\frac{1}{2\pi R}\int_0^{2\pi} u(s,\theta)R\, d\theta
%= \frac{1}{2\pi R}\int_0^{2\pi} \sum_{i,j} a_{i,j} \phi_{i,j}(s,\theta) R\, d\theta
%\\= \frac{1}{2\pi R}\sum_{i,j} a_{i,j}\int_0^{2\pi}  \phi_{i,j}(s,\theta) R\, d\theta
%=  \sum_{i} a_{i,0} \phi_{i}(s).
%\end{multline*}
%From \cite[Lemma 4.11]{chandler2015interpolation} we have
%\begin{equation}\label{H12norm_Gamma}
%\|u\|^2_{H^{\frac 12}(\Gamma)}=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left( 1+ \rho_{ij}\right)^{\frac 12}|a_{ij}|^2,
%\text{ with }
%a_{ij}=\int _0^1\int _0^{2\pi} u(s,\theta )\phi_{ij}\, R d\theta ds.
%\end{equation}
%and 
%\begin{equation*}
%\|\avrc{u}\|^2_{H^{\frac 12}(\Lambda)}=\sum_{i=1}^{\infty} \left( 1+ \rho_{i}\right)^{\frac 12}|\avrc{a}_i|^2,
%\text{ with }
%\avrc{a}_i=\int _0^1 \avrc{u}(s )\phi_{i}(s) ds.
%\end{equation*}
%
%Therefore, we have
%\begin{multline*}
%\|\avrc{u}\|^2_{H^{\frac 12}(\Lambda)}=
%\sum_{i=1}^{\infty}\left( 1+ i^2\pi^2\right)^{\frac 12}\left( \int_0^1 \avrc{u}(s) sin(i\pi s)\, ds \right)^2\\
%= \sum_{i=1}^{\infty} \left( 1+ i^2\pi^2\right)^{\frac 12}\left( \sum_{j=1}^\infty a_{j,0}\int_0^1 \sin(j\pi s) \sin(i\pi s) \, ds  \right)^2\\
%= \sum_{i=1}^{\infty} \frac 14 \left( 1+ i^2\pi^2\right)^{\frac 12}a_{i,0}^2\\
%\leq \sum_{i=1}^{\infty} \sum_{j=1}^{\infty}  \left( 1+ i^2\pi^2 + \frac{j^2}{R^2}\right)^{\frac 12} |a_{i,j}|^2 =\|u\|^2_{H^{\frac 12}(\Gamma)},
%\end{multline*}
%
%where we have used the fact that
%\begin{eqnarray*}
%&\int_0^1 \sin(i\pi s) \sin(j\pi s)\, ds=0 \quad \text{if $i\neq j$}\\
%&\int_0^1 \sin(i\pi s) \sin(j\pi s)\, ds=\frac 12 \quad \text{if $i =j$}.
%\end{eqnarray*}
%Moreover, in the case in which $u$ is constant on each cross section, from \eqref{H12norm_Gamma} we have
%\begin{multline*}
%\|u\|^2_{H^{\frac 12}(\Gamma)}=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left( 1+ \rho_{ij}\right)^{\frac 12}|a_{ij}|^2
%=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left(  1+ i\pi ^2+\frac{j^2}{R^2}\right)^{\frac 12}\left( \int _0^1\int _0^{2\pi} u(s,\theta )\phi_{ij}\, R d\theta ds \right)^2\\
%=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left(  1+ i\pi ^2+\frac{j^2}{R^2}\right)^{\frac 12}\left( \int _0^1 u(s) \int _0^{2\pi} \phi_{ij}\, R d\theta ds \right)^2,
%\end{multline*}
%and using \eqref{null_int_eigenf} and \eqref{nonull_int_eigenf}, we obtain
%\begin{multline*}
%\|u\|^2_{H^{\frac 12}(\Gamma)}=
%\sum_{i=1}^{\infty}\left( 1+ i\pi ^2\right)^{\frac 12}\left(\int _0^1 u(s )sin (i\pi s) 2\pi R ds\right)^2\\
%=4\pi ^2 R^2 \sum_{i=1}^{\infty}\left( 1+ \rho _i\right)^{\frac 12}|a_i|^2 = 4\pi ^2 R^2  \|u\|^2_{H^{\frac 12}(\Lambda)}.
%\end{multline*}
%\hspace*{0.9\textwidth} c.v.d.
%\end{proof}
%\begin{remark}
%The results of \eqref{lemma:H12norm}  can be generalized to the case of a  different geometry of $\Gamma$, for example a parallelepiped.  
%\end{remark}

\begin{lemma}\label{lemma:H12norm}
Let $\Gamma$ be a tensor product domain, $\Gamma= (0,X) \times (0,Y)$. For any regular $u(x,y)$ in $\Gamma$, let $\avrc{u}(x)=\frac 1Y \int _0^Y u(x,y)\, dy$. Then, for any $u\in H_{00}^{\frac 12}(\Gamma)$, then $\avrc{u}(x)\in H_{00}^{\frac 12}((0,X))$. 
Moreover, if $u(x,y)\in H^{\frac 12}_{00}(\Gamma)$ is constant with respect to $y$, namely $u(x,y)=u(x)$, then 
\begin{equation*}
\|u\|_{H^{\frac 12}_{00}(\Gamma)}=Y \|u\|_{H^{\frac 12}_{00}(0,X)}.
\end{equation*}
\end{lemma}
\begin{proof}
Let us consider the eigenvalue problems for the Laplace operator on $\Gamma$ with homogeneous Dirichlet conditions at $x=0,\, X$ and periodic boundary conditions at $y=0,\, Y$ and on $(0,X)$ with homogeneous Dirichlet conditions. Let us denote as $\phi _{ij}$ and $\rho _{ij}$, for $i=1,2,\dots$, $j=0,1,\dots$, the eigenfunctions and the eigenvalues of the laplacian on $\Gamma$, and with $\phi _i$ and $\rho _i$ the eigenfunctions and the eigenvalues of the laplacian on $(0,X)$. In particular,
\begin{align*}
\phi _{ij}(x,y)=\sin \left(\frac{i\pi x}{X}\right)\left( cos\left(\frac{j2\pi y}{Y}\right)+ sin	\left(\frac{j2\pi y}{Y}\right) \right),\\
\rho_{ij}=\left(\frac{i\pi}{X}\right) ^2+\left(\frac{j2\pi}{Y}\right)^2,\\
\phi _{i}(x)=\sin \left(\frac{i\pi x}{X}\right),\\
\rho _i = \left(\frac{i\pi}{X}\right) ^2.
\end{align*}
It is easy to verify that 
\begin{eqnarray}
\label{null_int_eigenf}
\int_0^{Y} \phi _{ij}(x,y)=0 \quad \forall j>0, \forall i \\
\label{nonull_int_eigenf}
\int_0^{Y} \phi _{ij}(x,y)= Y \, \sin\left(\frac{i\pi x}{X}\right) \quad \mbox{if } j=0, \forall i  .\\
\end{eqnarray}
Moreover we recall that $\phi_{i,j}(x,y)$ and $\phi _i(x)$ are orthogonal basis of $L^2(\Gamma)$ and $L^2(0,X)$ respectively. Therefore,
\begin{multline*}
\avrc{u}(x)=\frac{1}{Y}\int_0^{Y} u(x,y)\, dy
= \frac{1}{Y}\int_0^{Y} \sum_{i,j} a_{i,j} \phi_{i,j}(x,y) \, dy
\\= \frac{1}{Y}\sum_{i,j} a_{i,j}\int_0^{Y}  \phi_{i,j}(x,y) \, dy
=  \sum_{i} a_{i,0} \phi_{i}(x).
\end{multline*}
From \cite[Lemma 4.11]{chandler2015interpolation} we have
\begin{equation}\label{H12norm_Gamma}
\|u\|^2_{H^{\frac 12}(\Gamma)}=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left( 1+ \rho_{ij}\right)^{\frac 12}|a_{ij}|^2,
\text{ with }
a_{ij}=\int _0^X\int _0^{Y} u(x,y )\phi_{ij} (x,y)dy dx.
\end{equation}
and 
\begin{equation*}
\|\avrc{u}\|^2_{H^{\frac 12}(0,X)}=\sum_{i=1}^{\infty} \left( 1+ \rho_{i}\right)^{\frac 12}|\avrc{a}_i|^2,
\text{ with }
\avrc{a}_i=\int _0^X \avrc{u}(x )\phi_{i}(x) dx.
\end{equation*}

Therefore, we have
\begin{multline*}
\|\avrc{u}\|^2_{H^{\frac 12}(0,X)}=
\sum_{i=1}^{\infty}\left( 1+ \left(\frac{i\pi}{X}\right)^2\right)^{\frac 12}\left( \int_0^X \avrc{u}(x) \sin\left(\frac{i\pi x}{X}\right)\, dx \right)^2\\
= \sum_{i=1}^{\infty} \left( 1+ \left(\frac{i\pi}{X}\right)^2\right)^{\frac 12}\left( \sum_{j=1}^\infty a_{j,0}\int_0^1 \sin\left(\frac{j\pi x}{X}\right) \sin\left(\frac{i\pi x}{X}\right) \, dx  \right)^2\\
= \sum_{i=1}^{\infty} \frac{X^2}{4} \left( 1+ \left(\frac{i\pi}{X}\right)^2\right)^{\frac 12}a_{i,0}^2\\
\leq \frac{X^2}{4}\sum_{i=1}^{\infty} \sum_{j=1}^{\infty}  \left( 1+ \left(\frac{i\pi}{X}\right)^2 + \left(\frac{j2\pi}{Y}\right)^2\right)^{\frac 12} |a_{i,j}|^2 =\frac{X^2}{4}\|u\|^2_{H^{\frac 12}(\Gamma)},
\end{multline*}

where we have used the fact that
\begin{eqnarray*}
&\int_0^X \sin\left(\frac{i\pi x}{X}\right) \sin\left(\frac{j\pi x}{X}\right)\, dx=0 \quad \text{if $i\neq j$}\\
&\int_0^X \sin\left(\frac{i\pi x}{X}\right) \sin\left(\frac{j\pi x}{X}\right)\, dx=\frac X2 \quad \text{if $i =j$}.
\end{eqnarray*}
Moreover, in the case in which $u$ is constant with respect to $y$, from \eqref{H12norm_Gamma} we have
\begin{multline*}
\|u\|^2_{H^{\frac 12}(\Gamma)}=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left( 1+ \rho_{ij}\right)^{\frac 12}|a_{ij}|^2\\
=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left(  1+ \left(\frac{i\pi}{X}\right)^2 + \left(\frac{j2\pi}{Y}\right)^2\right)^{\frac 12}\left( \int _0^X\int _0^Y u(x,y )\phi_{ij}(x,y) \,dy\,dx \right)^2\\
=\sum_{i=1}^{\infty}\sum_{j=0}^{\infty} \left(  1+ \left(\frac{i\pi}{X}\right)^2 + \left(\frac{j2\pi}{Y}\right)^2\right)^{\frac 12}\left( \int _0^X u(x) \int _0^Y \phi_{ij}(x,y)\,  dy dx \right)^2,
\end{multline*}
and using \eqref{null_int_eigenf} and \eqref{nonull_int_eigenf}, we obtain
\begin{multline*}
\|u\|^2_{H^{\frac 12}(\Gamma)}=
\sum_{i=1}^{\infty}\left( 1+ \left(\frac{i\pi}{X}\right)^2\right)^{\frac 12}\left(\int _0^X Yu(x)\, \sin\left(\frac{i\pi x}{X}\right) dx\right)^2\\
=Y^2 \sum_{i=1}^{\infty}\left( 1+ \rho _i\right)^{\frac 12}|a_i|^2 = Y^2  \|u\|^2_{H^{\frac 12}(0,X)}.
\end{multline*}
\end{proof}

\begin{corollary}
Let $\Gamma$ be the lateral surface of a prism with $N$ faces $\Gamma_i$, $i=1, \dots, N$, with $\Gamma_i=(0,X_i)\times(0,S)$. For any $u\in H^{\frac 12}_{00}(\Gamma)$, let $\avrc{u}(s)= \frac{1}{\sum_i{|X_i|}}\sum_{i=1}^N \int_0^{X_i} u(x,s)\, dx$. Then $\avrc{u}\in H^{\frac 12}_{00}(0,S)$.
\end{corollary}

\begin{corollary}
Let $\Gamma$ be the lateral surface of a cylinder of radius $R$. Let $u\in H^{\frac 12}_{00}(\Gamma)$ and let $\avrc{u}(s)= \frac{1}{2\pi R} \int_0^{2\pi} u(s,\theta)R\, d\theta$. Then $\avrc{u}\in H^{\frac 12}_{00}(0,S)$.
\end{corollary}

It is apparent that problems \eqref{eq:weak_dirneu} and \eqref{eq:red_dirneu} share the same mathematical structure.
For this reason, the well-posedness of \eqref{eq:red_dirneu} can be studied in the framework of the classical theory of saddle point problems.

%Assuming now that 
%$\Pi_1: H^1(\Omega) \rightarrow H^{\frac12}(\Gamma)$ or $\Pi_2: H^1(\Gamma) \rightarrow H^{\frac12}(\Gamma)$ are isomorphisms in the sense that 
%\[
%\| \Pi_1 \|_{L(H^1(\Omega), H^{\frac12}(\Gamma))} \le C \mbox{, } 
%\| \Pi_1^{-1} \|_{L(H^{\frac12}(\Gamma), H^1(\Omega))} \le 1/c,  
%\]
%and 
%\[
%\| \Pi_2 \|_{L(H^1(\Omega), H^{\frac12}(\Gamma))} \le C \mbox{, } 
%\| \Pi_2^{-1} \|_{L(H^{\frac12}(\Gamma), H^1(\Omega))} \le 1/c .   
%\]
%We obtain the inf-sup condition as follows. Assume that $\Pi_1$ is an isomorphism then  
%we may show the inf-sup conditioning by letting $\Pi_1 \hat u = M $ and $ \hat \ud = 0$ (i.e., we need the lower bound only for one
%of the operators).  
%\begin{eqnarray}
%\sup_{u,\ud}\frac{
%\langle \Pi_1 u - \Pi_2 \ud , M \rangle_{(H^{\frac12}(\Gamma), H^{-\frac12}(\Gamma))} 
%}{(|u|_1^2 + |\ud|_1^2)^{1/2}} 
%\ge \\ 
%\frac{ \langle \Pi_1 \hat u - \Pi_2 \hat \ud , M \rangle_{(H^{\frac12}(\Gamma), H^{-\frac12}(\Gamma))} 
%}{(|\hat u|_1^2 + |\hat \ud|_1^2)^{1/2}}  \ge \\
%\frac{ \langle M , M \rangle_{(H^{\frac12}(\Gamma), H^{-\frac12}(\Gamma))} 
%}{|\Pi_1^{-1} M |_{H^1(\Omega})} \ge \\ 
%\frac{ \langle M , M \rangle_{(H^{\frac12}(\Gamma), H^{-\frac12}(\Gamma))} 
%}{| M |_{H^\frac12(\Gamma)}} \ge \\ 
%c | M |_{H^{-\frac12}(\Gamma)}  
%\end{eqnarray}
%I think the other Brezzi conditions also follows, but for the boundedness we
%would need that both $\Pi_1$ and $\Pi_2$ are bounded. Not sure I am 
%happy with the $\langle \cdot, \cdot \rangle$ notation and in particular not when we spesify the
%spaces. I would rather just say that $(\cdot, \cdot)$ denotes both the $L_2$ inner product and the
%duality pairing.  
%If we find it useful we may also  
%consider $\Pi_2: H^1(\Gamma) \rightarrow X(\Gamma)$ where $X$ may e.g. be $H^1$.  
%If so, $L, M$ in $ H^{-\frac12}(\Gamma) \cap X(\Gamma)$

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%Analysis of the continuous problem 
\input{infsup_continuous}

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%Analysis of the discrete problem in the case in which the LM lives on the coarser mesh w.r.t the trace mesh deriving from the 3D mesh

%\input{infsup_discrete_coarserLM}

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%Analysis of the discrete problem in the case in which the LM lives on the the trace mesh deriving from the 3D mesh (conforming to the interface cylinder \Gamma)

\input{infsup_discrete_conformingLM}

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%Analysis of the discrete problem in the case in which we have a non conforming 3D mesh (just in the case of formulation 2 3D-1D-1D)

\input{unfitted_prob2}

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

\input{benchmark}
%-----------------
\bibliographystyle{siamplain}
\bibliography{3d1d_coupled}
\end{document}